---
title: "SQL Practice_Olympic Data"
output: html_document
---

```{r}
library(DBI)
```

```{r}
install.packages("dplyr", "RSQLite")
```

```{r}
library(tidyverse)
library(dplyr)
library(dbplyr)
library(readr)
library(here)
```

```{r}
con <- dbConnect(RSQLite::SQLite(), dbname = ":memory:")
dbWriteTable(con, "summer_medals", read.csv("/Users/joypham/LearningData/LearningR/olympics/summer.csv"))
```

## Get first 6 rows of the file
```{r}
summer_medals <- read_csv("/Users/joypham/LearningData/LearningR/olympics/summer.csv", col_names = TRUE)

head(summer_medals)
```

 A quick sneak peak a the file structure
 
```{r}
str(summer_medals)
```

Get the distinct disciplines

```{sql connection=con}
SELECT DISTINCT discipline
FROM summer_medals
```
## Gold in volleyball

Volleyball made its first appeared as an Olympic game in 1964.

First, a "stupid" query, given that I didn't know how many members in a volleyball team.

```{sql connection=con}
 WITH sub AS (
  SELECT year,
       country,
       gender,
       COUNT(*) OVER(PARTITION BY country, gender ORDER BY year) AS winning_times
  FROM summer_medals
  WHERE discipline = "Volleyball" AND
      medal = "Gold"
  GROUP BY year, country, gender)
  
SELECT country, gender,
       MAX(winning_times) AS total_winning
FROM sub
GROUP BY country, Gender
       
```
Then, a more elegant and simpler query

```{sql connection=con}
 WITH sub AS(
  SELECT  country, gender,
         COUNT(*)/12 AS gold_count
  FROM summer_medals
  WHERE discipline = "Volleyball" AND
      medal = "Gold"
  GROUP BY country, gender
  ORDER BY gold_count DESC, country ASC)
 
 SELECT country, gender, SUM(gold_count) AS winning_times
 FROM sub
 GROUP BY country, gender
```

## Medals in baseball

Who got the first prize in baseball?

```{sql connection=con}
WITH sub AS (
  SELECT country,
         COUNT(*)/20 AS gold_count
  FROM summer_medals
  WHERE discipline = "Baseball" AND medal = "Gold"
  GROUP BY country
)

SELECT country,
       SUM(gold_count) AS winning_times
FROM sub
GROUP BY country
```

More details about the countries that were able to snatch a medal in baseball in each Olympic in 1992-2008

```{sql connection=con}

  SELECT year,
         country,
         medal,
         COUNT(*) OVER(PARTITION BY medal, country
                  ORDER BY year ASC) AS times
  FROM summer_medals
  WHERE discipline = "Baseball"
  GROUP BY year, country, medal
  ORDER BY year ASC

```
## Swimming giants
Who ruled the blue races?
Ladies first

```{sql connection=con}
  WITH sub AS (
    SELECT
         country,
         COUNT(medal) AS total_gold
  FROM summer_medals
  WHERE discipline = "Swimming" AND
        medal = "Gold" AND
        gender = "Women"
  GROUP BY country, Gender) 
  
SELECT country, total_gold,
       DENSE_RANK() OVER(ORDER BY total_gold DESC) AS ranking
FROM sub
LIMIT 10
```
Then, gentlemen

```{sql connection=con}
  WITH sub AS (
    SELECT
         country,
         COUNT(medal) AS total_gold
  FROM summer_medals
  WHERE discipline = "Swimming" AND
        medal = "Gold" AND
        gender = "Men"
  GROUP BY country, Gender) 
  
SELECT country, total_gold,
       DENSE_RANK() OVER(ORDER BY total_gold DESC) AS ranking
FROM sub
LIMIT 10
```

## Assigning row number to the data set

```{sql connection=con}
SELECT 
*,
-- Assign numbers to each row
ROW_NUMBER() OVER() AS Row_N 
FROM summer_medals
ORDER BY Row_N ASC;
```

## Numbering Olympic games in ascending order

```{sql connection=con}
SELECT
  year,
  ROW_NUMBER() OVER() AS Row_N 
FROM (
  SELECT DISTINCT year
  FROM summer_medals
  ORDER BY year ASC
) AS year
ORDER BY year ASC;
```

## Find the total medals each year
Note: SQLite cannot perform `ROLLUP()`

```{sql connection=con}
SELECT year, 
       medal,
       COUNT(*) AS total
FROM summer_medals
GROUP BY year, medal
ORDER BY year
```
##Ranking Olympic athletes by medals

```{sql connection=con}
WITH sub AS (
  SELECT athlete,
         country,
         discipline,
         COUNT(*) AS medals
  FROM summer_medals
  GROUP BY athlete
  ORDER BY medals DESC
)

SELECT athlete,
       country,
       discipline,
       medals,
       DENSE_RANK() OVER(ORDER BY medals DESC) AS medals_rank
FROM sub
ORDER BY medals_rank;
```
##Reigning 1500m Freestyle champions

```{sql connection=con}
WITH champions AS (
  SELECT 
    athlete AS champion,
    year,
    country
  FROM summer_medals
  WHERE discipline = "Swimming"
  AND
    event = "1500M Freestyle" AND
    gender = "Men" AND
    medal = "Gold"
)

SELECT year,
       country,
       champion,
       LAG(champion, 1) OVER(ORDER BY year ASC) AS last_champion
FROM champions
ORDER BY year ASC;
```
## Marathon and high jump after the Cold War

```{sql connection=con}
 
  SELECT event, country, gender,
         COUNT(*) AS gold_count
  FROM summer_medals
  WHERE year > 1964 AND
        discipline = "Athletics" AND
        event IN ("Marathon", "High Jump") AND
        medal = "Gold"
 GROUP BY event, country, gender
 ORDER BY gold_count DESC, country ASC
  

```

## Next city in line

```{sql connection=con}
WITH sub AS (
  SELECT year, city
  FROM summer_medals
  GROUP BY year, city
)

SELECT year, city,
       LEAD(city,1) OVER(ORDER BY year) AS next_city
FROM sub
ORDER BY year ASC
```

## Last Judo champion after Cold War (- 60 KG)

```{sql connection=con}
WITH sub AS (
  SELECT year, country
  FROM summer_medals
  WHERE discipline = "Judo" AND
        event = "- 60 KG" AND
        gender = "Men" AND
        medal = "Gold")
        
SELECT year, country,
       LAST_VALUE(country) OVER(
         ORDER BY year ASC
         RANGE BETWEEN UNBOUNDED PRECEDING AND UNBOUNDED FOLLOWING) AS last_champion
FROM sub
ORDER BY year ASC
```

## Ranking athletes from Japan, China, and Korea

```{sql connection=con}
WITH sub AS (
  SELECT country, athlete, discipline, COUNT(*) AS medals
  FROM summer_medals
  WHERE country IN ("CHN", "JPN", "KOR") AND
        year >= 1984 AND medal = "Gold"
  GROUP BY country, athlete, discipline
  HAVING COUNT(*) > 1
)

SELECT country, athlete, discipline, medals,
       DENSE_RANK() OVER(PARTITION BY country ORDER BY medals DESC) AS ranking
FROM sub
ORDER by ranking ASC
```
## Total medals from 1984 (JPN, KOR, CHN)

Medals earned by year

```{sql connection=con}
  SELECT year, country,
         medal,
         COUNT(*) AS medal_count
  FROM summer_medals
  WHERE country IN ("JPN", "KOR", "CHN") AND
        year >= 1984
  GROUP BY year, country, medal
  ORDER BY year ASC

```
Total earned medals from 1984

```{sql connection=con}
  WITH sub AS(
  SELECT year, country,
         medal,
         COUNT(*) AS medal_count
  FROM summer_medals
  WHERE country IN ("JPN", "KOR", "CHN") AND
        year >= 1984
  GROUP BY year, country, medal
  ORDER BY year ASC)

SELECT country, medal,
       SUM(medal_count) AS total
FROM sub
GROUP BY country, medal
```


```{sql connection=con}
WITH sub1 AS(
  SELECT year,
         COUNT(*) AS gold_count
  FROM summer_medals
  WHERE country IN ("JPN", "KOR", "CHN") 
        AND medal = "Gold"
        AND year >= 1984
  GROUP BY year
),
sub2 AS (
  SELECT year, COUNT(*) AS total
  FROM summer_medals
  WHERE medal = "Gold" AND year >=1984
  GROUP BY year
  ORDER BY year ASC
)

  SELECT sub1.year, sub1.gold_count, sub2.total
  FROM sub1
  INNER JOIN sub2
  ON sub1.year = sub2.year
  ORDER BY sub1.year ASC

```

